\documentclass{article}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{../mystuff}
\usepackage{apalike}
\title{WAFO Data Structures: Wave Spectrum and Covariance Function}
\author{ Eva Sj\"o}
\begin{document}
\maketitle
\bibliographystyle{apalike}
\section{Introduction}
The following types of wave spectra are available in WAFO:
\begin{itemize}
\item Directional spectrum (at fixed position). Defining $Cov\lbrace {\eta}(0,0,0),{\eta}(t,x,y)\rbrace$.\\
\item Frequency spectrum (at fixed position). Defining $Cov\lbrace {\eta}(0){\eta}(t)\rbrace$.\\
\item Wave number spectrum in $2$-D. Defining $Cov\lbrace {\eta}(0,0,0),{\eta}(t,x,y)\rbrace$.\\
\item Wave number spectrum in $1$-D. Defining $Cov\lbrace{\eta}(0),{\eta}(x)\rbrace$.\\
\item Encountered directional spectrum (for given speed and
  direction). Defining $Cov\lbrace {\eta}(0,0,0), {\eta}(t,x+vt\cos{\phi},y+vt\sin{\phi})\rbrace$. \\
\item Encountered frequency spectrum (for given speed and
  direction). Defining $Cov\lbrace {\eta}(0,0,0), {\eta}(t,vt\cos{\phi},vt\sin{\phi})\rbrace$.\\
\item Rotated directional spectrum. Defining $\Cov
  \lbrace{\eta}_{\phi}(0,0,0),{\eta}_{\phi}(t,x,y)\rbrace$, ${\eta}_{\phi}(\cdot)$=wave field
  rotated angle ${\phi}$.\\
\item Rotated wave number spectrum in 2-D. Defining
  $Cov\lbrace{\eta}(0,0,0),{\eta}_{\phi}(t,x,y)\rbrace$.\\
\item Rotated wave number spectrum in 1-D. Defining
  $Cov\lbrace{\eta}(0,0,0),{\eta}(0,x\cos{\phi},x\sin{\phi})\rbrace$.
\end{itemize}

The idea of all these spectra is quite clear, but there are several possible
definitions for each of them when it comes to units, interval of
definition of parameters etc.

The suggestion here is to allow for different frequency units, but try to keep to only one definition
for each type of spectrum (apart from the units).

Further, the spectra should not be assumed symmetric in
${\omega}$. This we can call the physical definition. The mathematical 
definition with (conjugate) symmetry will be used  when needed internal in the routines.

With $g=$constant of gravity$\approx 9.81$ and ${\omega}=2{\pi}f$ we have
\begin{description}
\item[Directional spectrum] $S({\omega},{\theta})$, $0 <
  {\omega}<\infty$, $-{\pi}<{\theta}<{\pi}$. The main direction of
  waves is ${\theta}=0$.
  Alternatively we
  have $S_f(f,{\theta})$, $0 <
  f<\infty$, $-{\pi}<{\theta}<{\pi}$. Remark: The definition used
  in \cite{Lindgren-Rychlik-Prevosto_97b} is  $-\infty <
  {\omega}<\infty$, $-{\pi/2}<{\theta}<{\pi/2}$ with the motivation
  that it gives more transparent formulas for the calculation of
  encountered spectra.
\item[Frequency spectrum] $S({\omega})$, $0 \leq {\omega} < \infty, 
  $ or $S_f(f)$, $0 < f[Hz] < \infty$. 
  Can be integrated out from the directional spectrum:  $S({\omega})=\int_{-\pi}^{\pi}S({\omega},{\theta})d{\theta}$.
  \\
\item[Wave number spectrum in 2-D]
  $S_k({\kappa}_x,{\kappa}_y)=S_k(\mbf{\kappa})$, $-\infty <{\kappa}_x,{\kappa}_y<\infty$.  Related 
  to the directional spectrum via
\begin{xalignat*}{2}
  \label{dr_param}
  &   \begin{cases}
    {\omega}(\mbf{\kappa})=\sqrt{g||\mbf{\kappa}||\tanh{(||\mbf{\kappa}||h)}}\\
    {\theta}(\mbf{\kappa})=\arctan_2({\kappa}_y/{\kappa}_x).
  \end{cases}&\text{or}\quad &\begin{cases}
  {\kappa}_x({\omega},{\theta})=k({\omega};h)\cos{\theta} \\
  {\kappa}_y({\omega},{\theta})=k({\omega};h)\sin{\theta}
 \end{cases}
\end{xalignat*}
where ${\kappa}({\omega};h)$ is the solution to the dispersion
relation ${\omega}^2=g{\kappa}\tanh({\kappa}h)$, the dispersion
relation for water depth $h$. The notation $\arctan_2$ means the four
quadrant arctangent, and $||\mbf{\kappa}||=\sqrt{{\kappa}_x^2+{\kappa}_y^2}$. For deep water ($h=\infty$)  we have
$S_k({\kappa}_x,{\kappa}_y)=\frac{g^2}{2{\omega}(\mbf{\kappa})^3}S({\omega}(\mbf{\kappa}),{\theta}(\mbf{\kappa}))$.
Maple helped me to evaluate the Jacobian determinant for the change of 
variables from $\mbf{\kappa}$ to $({\omega},{\theta})$ for finite depth, but I do not
think you would like to see that expression  here...
Similar relation can be derived to $S_f(f,{\theta})$ (involving more $2{\pi}$'s\ldots)\\
\item[Wave number spectrum in 1-D] $S_k({\kappa})$, $0\leq {\kappa} < \infty$. For deep water related to $S({\omega})$
  through $S(w)=\frac{2{\omega}}{g}S_k({\omega}^2/g)$. Or reversely, $S_k({\kappa})=\frac{1}{2}\sqrt{\frac{g}{{\kappa}}}S(\sqrt{g{\kappa}})$. For finite depth the
  dispersion relation is given by
  ${\omega}^2=g{\kappa}\tanh({\kappa}h)$, and from this a relation can 
  be found.\\
\item[Encountered frequency spectrum] $S_e({\omega};v,{\phi})$, for ships traveling with speed 
  $v$  in the direction ${\phi}$.
  See~\cite{Lindgren-Rychlik-Prevosto_97b} and
  ~\cite{Podgorski-Rychlik-Machado} how it
  is derived from a directional spectrum.
\item[Encountered directional spectrum] $S({\omega},{\theta};
  v,{\phi})$. Same as previos, but without integrating out the angle.\\  
\item[Rotated directional spectrum]
  $S_{\phi}({\omega},{\theta})=S({\omega},{\theta}+{\phi})$, where
  $S({\omega},{\theta})$ is extended periodically for ${\theta}$
  beyond $(-{\pi},{\pi})$.
\item[Rotated wave number spectrum] $S_k({\kappa};{\phi})$, spectrum
  along a line in an angle ${\phi}$ with the main direction of waves. In 2-D: $S_k({\kappa}_x,{\kappa}_y;{\phi})$, $x$-axis
  in an angle ${\phi}$ with the main direction of waves. Can be
  derived from $S_{\phi}({\omega},{\theta})$.
\end{description}

\section{In MATLAB}
All this is possible to describe in MATLAB with a so called 'Structured 
Array'. The fields needed to describe the spectra defined above are
\begin{description}
\item[.type] Defining type of spectrum: 'dir', 'freq', 'k2D', 'k1D',
  'encdir', 'enc', 'rotdir', 'rotk2D', 'rotk1D'. Default: 'freq'.\\
\item[.S] Values of spectrum in a matrix size $n_{\theta} \times
  n_f$. 
\item[.w] Frequency lag when angular frequency, vector of length
  $n_f$.  Equidistant values $\geq 0$. Should not be given when .f or .k (see
  below) is. Default: see .S.\\
\item[.f] Frequency lag when natural frequency ([Hz]), vector of length
  $n_f$. Equidistant values $\geq 0$. Should not be given when .w or .k (see
  below) is. Default: see .S.\\
\item[.k] Wave number lag (in first dimension), vector of length
  $n_f$.  Equidistant values, $\geq 0$ when type 'k1D', negative
  values allowed when type 'k2D' etc. Only when some type of wave number
  spectrum. Default: see .S.\\
\item[.k2] Only for .type='k2D' or 'rotk2D'. Second dimension wave number
  lag, length $n_{\theta}$, equidistant. Default: equal to .k.\\
\item[.theta] Only for all types of dirctional spectra. Angular lags
  $-{\pi}<{\theta}[\text{rad}]<{\pi}$,
   equidistant vector of length $n_{\theta}$.  Default: see .S\\
\item[.v] Only for .type='encdir' and 'enc'. Speed of ship, scalar,
  unit [m/s]. Default: 0\\
\item[.phi] Only for .type='encdir', 'enc' and 'rot\ldots'. Direction of ship 
  or direction of '$x$-axis', scalar, unit [rad]. Defualt: 0\\
\item[.h] Water depth, positive scalar, unit [m]. Default: $\infty$\\
\item[.tr] Transformation funtion. Default: [] (none)
\item[.norm] Normalization flag. Logical 1 if variance normalized
  spectrum, 0 else. Default: 0
\item[.note] String for memorandum about the spectrum.
\item[.date] Date and time of creation or change of the spectrum.
\end{description}
Remark: The field .type in in most cases given by the other fields in 
the sense that the combination of fields, and the values in these,
uniquely describes the .type.  For example, if .v is set, then we have a 
encountered spectrum of some kind, if .theta also is set, then we have
an encountered directional spectrum. A reason to keep the field .type anyway
is that it provides an easy check of type of spectrum.
Remark 2: The fields .w, .f and .k are disjoint.

The following routines are closly related to the spectrum structure:
%\textbf{findtype(spec)} that looks at the fields
%specified in variable spec at decides which .type it must
%be.
\textbf{checkspec(spec)} to check if  spec.type is compatible with the
rest of the fields in spec, or if any field has to be changed.\\
\textbf{spec2spec(spec,newtype)} that can
translate from one type to another when possible, see
Table~\ref{tab:changes}.\\
\textbf{createspec(type, freqtype)} that sets up an empty structure.
\textbf{w2k} and \textbf{k2w} translates between the frequency world
and the wave number world in both one and two dimensions.
\begin{table}[htbp]
  \begin{center}
\begin{tabular}[h]{|c||c|c|c|c|c|c|c|c|c|}
\hline .type & dir & freq & k2D & k1D & encdir & enc & rotdir &
rotk2D & rotk1D\\\hline\hline
dir & $\bullet$ &$\Rightarrow$&  $\Rightarrow$ &  $\rightarrow$ &
$\Rightarrow$ &  $\Rightarrow$ &  $\Rightarrow$ &  $\Rightarrow$ &
$\rightarrow$ \\ \hline
freq & & $\bullet$ & &  $\Rightarrow$ & & & &&\\\hline
k2D &  $\Rightarrow$ &  $\rightarrow$ &  $\bullet$ &
$\Rightarrow$ &  $\rightarrow$ & $\rightarrow$ &  $\rightarrow$ &
$\Rightarrow$ &  $\rightarrow$ \\\hline
k1D & &  $\Rightarrow$ && $\bullet$ &&?&&&\\\hline
encdir & ? & ? & ? & ? &  $\bullet$ &  $\Rightarrow$ & ? & ? & ?
\\\hline
enc & & ? & & & & $\bullet$ & & & ?\\\hline
rotdir &  $\Rightarrow$ &  $\Rightarrow$ &  $\rightarrow$ &
$\rightarrow$ &  $\rightarrow$ &  $\rightarrow$ &  $\bullet$ &
$\Rightarrow$ & $\Rightarrow$ \\\hline
rotk2D&  $\rightarrow$ &  $\rightarrow$ &  $\Rightarrow$ &
$\rightarrow$ &  $\rightarrow$ & $\rightarrow$ &  $\Rightarrow$ &
$\bullet$ &  $\Rightarrow$ \\\hline
rotk1D & &  $\Rightarrow$ &&  &&?&&& $\bullet$ \\\hline
\end{tabular}
    \caption{Scheme of possible type changes. An empty box means
      translation impossible,  $\Rightarrow$ means
      direct translation, $\rightarrow$ means translaition via another 
      type is most easy, and ? means that I had no time to find out if it is possible. }
    \label{tab:changes}
  \end{center}
\end{table}

NB! All field names and file names above are only suggestions, any
idea of improvement is welcome, as well as comments/corrections on the 
Table.
\section{Covariance function data structure}
To keep everything gathered, the covariance function is also stored in 
a structured array together with its derivatives and other thing that
can be useful. One idea with this structure is that although many
fields are included in the list below, giving possibilities to cover a
wide range of different situations, it is preferable to only include  the
fields actually needed. For example, we allow for any combination of the varables $x,y$ and
$t$, and for up to four derivatives, but if we have the covariance
function $R(t)$, then no $x$- or $y$-variables or -derivatives should be in
the struct for this cvf. When we have more than one variable, the
mutual  order in the
struct between them should always be $x,y,t$.

This may seem complicated, but to make it simpel there is a
script:\\\textbf{createcov(nrofder,variables,type)} that sets up an
empty struct with the fields given by the options: nrofder=number of 
derivatives, variables= a string of up to 3 characters of any
combination of x,y and t, type='none','enc' or 'rot'. All the possible 
types for spectra are not interesting for covariance functions,
but if we have \textsl{any} of the rotated types or \textsl{any} of the encountered
types are relevant also for cvfs, all other types passes as 'none'.

Here follows a list of all possible fields (if I am right, worst case
is 44 fields...) in recommended order:
\begin{description}
\item[.R] Matrix/vector with covariance function. Size $n_x \times
  n_y\times  n_t$, but no singleton dimesions. If no y, for example,
  then the size is $n_x \times n_t$.
\item[.x] Lag of first space dimension, vector of length $n_x$, unit
  [m].
\item[.y] Lag of second space dimension, vector of length $n_t$, unit
  [m].
\item[.t] Time lag, vector of length $n_t$, unit [s].
%\item[.stdev] Only when one-dimensional (time) function.
\item[.h] Water depth, scalar, unit [m].
\item[.tr] Transformation, see spectrum structure.
\item[.type] 'enc','rot' or 'none', see above.
\item[.v] Only if type 'enc'. Speed of ship, scalar, unit [m/s].
\item[.phi] Only if type 'enc' or 'rot'. Direction of ship 
  or direction of '$x$-axis', scalar, unit [rad].
\item[.norm] Normalization flag. Logical 1 if autocorrelation function(=
  normalized cvf), 0 else.
\item[.Rx \ldots .Rtttt] Matrices with obvious derivatives of R.  
\item[.note] String for memorandum about the cvf.
\item[.date] Date and time of creation or change of the cvf.
\end{description}
\bibliography{../mybib}

\end{document}
