%\documentclass[a4paper,11pt,oneside]{report}  % Draft settings
\documentclass[a4paper,11pt]{book}  % 11pt in final form

\newcommand{\FileIdX}{File: lab.tex, Last changed: 2005-04-14}
\newcommand{\FileId}{\FileIdX}
\newcommand{\IsDraft}{0} % 0 = Final version, 1 = Draft version


% Include Only

%\includeonly{lab1,lab2}
%\includeonly{lab2,lab3,lab4}
%\includeonly{flurp}

% Packages
\usepackage[T1]{fontenc}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{latexsym}
\usepackage{rotating}
\usepackage{fancyheadings}
\usepackage{amssymb}
\usepackage{ifthen}
\usepackage{verbatim}

% Local pacages
\usepackage{litenfig}

% Style-file for LAB
\usepackage{lab}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Set Font for LAB
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{times}             % Times
\usepackage{palatino}          % Palatino
%\renewcommand{\rmdefault}{pad}  % Garamond
%\renewcommand{\rmdefault}{pfr}  % Frutiger

\begin{document}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LAB
% - Title page
% - Chapters
% - Bibliography
% - Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\thispagestyle{empty}

\begin{center}
  {\bf {\Huge Random Loads with Data Analysis\\[2mm]
               {\LARGE --- Computer Exercises ---}}
  \ifthenelse{\equal{\IsDraft}{1}}{%
    \footnotetext{\FileId \hfill Compiled: \today}}{}

\vskip1cm

\begin{tabular}[t]{c@{\qquad}c}
\emph{Pär Johannesson}      & \emph{Igor Rychlik} \\[.5mm]
Fraunhofer-Chalmers Centre  & Mathematical Statistics \\
                            & Chalmers University of Technology \\[.5mm]
\end{tabular}

\vskip5mm

\ifthenelse{\equal{\IsDraft}{1}}{Draft: 2005-04-14}{Version
2005-04-14} }
\end{center}

\vskip5mm

%
% Table of Contents
%

%\tableofcontents

This short course is based on the course  ``Load and Fatigue
Analysis'' developed by Pär Johannesson, Igor Rychlik, Georg
Lindgren, and Jesper Ryd\'{e}n at Mathematical Statistics, Lund
Institute of Technology, where it was given in January 2000.  The
goal of the the course is to demonstrate the use of the Matlab
toolbox WAFO, Wave Analysis for Fatigue and Oceanography
(\verb+www.maths.lth.se/matstat/wafo/+), in the areas of fatigue
analysis and modelling of random loads.

The course took place at Fraunhofer-Chalmers Centre at Chalmers
Science Park with lectures on Wednesdays 15-17 and computer
exercises on Thursdays 10-12 March 2 and 3, April 6, 7, 13 and 14,
2005.


%\def\labelenumi{\theenumi.}        \def\theenumi{\arabic{enumi}}
\begin{enumerate}
  \setcounter{enumi}{-1}
  \item \textbf{General Introduction.} Introducing some basic concepts used
    in the following analysis.

  \item \textbf{Analysis of Load Data.}
    Analyse a load signal by means of turning points, rainflow
  filter, level crossings, irregularity, rainflow cycles, load
  spectrum, Palmgren-Miner damage, upper and lower bounds for
  rainflow damage. Fatigue life evaluation for constant and
  variable amplitude loads.
  \item \textbf{Markov modelling of loads.}
    A Markov chain of turning points is used for the modelling.
    The limiting rainflow matrix can be computed, and sample paths
    simulated.  The concept of switching loads is combined with
    the Markov model, and decomposition of a mixed rainflow matrix is treated.
  \item \textbf{Spectral modelling of loads.}
    A load is specified by its power spectrum. Its limiting
    rainflow matrix is computed through Markov approximation.
    Further, the expected damage can be computed using
    the narrow band approximation, or accurate numerical approximations.
  \item \textbf{Analysis of Measured Loads.}
    Analysis of a measured switching load.
    Estimation from measured time signal and from observed
    rainflow matrix. Decomposition of the mixed rainflow matrix
    into stationary blocks.
\end{enumerate}
%\def\labelenumi{\theenumi)}        \def\theenumi{\alph{enumi}}

%
% Login on the station
%

\newpage

\section*{Computers}

The students are supposed to bring their own laptops equipped with
Matlab and WAFO (Download: \verb+www.maths.lth.se/matstat/wafo/+).

%
% Starting Matlab
%

\subsection*{Starting Matlab with WAFO}
In order to access the routines needed for the exercises and start
Matlab, and add the path to the WAFO-root directory.
At the Matlab prompt, type
\begin{code}
>> initwafo('full')  % Initiate WAFO toolbox
>> itmlab            % Initiate Course files
\end{code}
Use the command \verb|help| whenever you need, e.g.\
\begin{code}
>> help
>> help wafo       % Contents of WAFO toolbox
>> help initwafo   % Help about initwafo
>> help fatigue    % Help about the fatigue part of WAFO
>> help dat2tp     % Help about routine dat2tp
>> help tp2rfc     % Help about routine tp2rfc
\end{code}


% ===================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cleardoublepage
\pagestyle{fancyplain}
%\pagestyle{headings}
%\pagestyle{fancy}

%\renewcommand{\chaptermark}[1]%
%  {\markboth{\textsl{Computer Exercise \thechapter. #1}}{}}

\include{lab_intro}
\include{lab1}
\include{lab2}
\include{lab3}
\include{lab4}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%\part{Appendix}
\cleardoublepage
\appendix

\renewcommand{\chaptermark}[1]%
  {\markboth{\textsl{Appendix \thechapter. #1}}{}}

%\include{wafoAppX}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cleardoublepage

\renewcommand{\FileId}{\FileIdX}

%\addcontentsline{toc}{chapter}{Bibliography}

\bibliographystyle{abbrv}
%\nocite{Johannesson97.M1,Johannesson97.R2,Johannesson97.Lic,Johannesson98.A1,Johannesson98.C2}
%\nocite{Lindgren91.A1,Rychlik97.A1,Rychlik92.A2,Rychlik93.A1,
%  Ochi94.A1,
%  Krenk89.A1}
%\bibliography{/usr/matstat/pj/tex/inputs/ref}
\bibliography{lab}

\end{document}
